# Regex - A quick introduction
A __Regular Expression__, or __regex__ for short, is a powerful tool, which helps us writing code for pattern matching in texts.  

## Literal Characters

Every _literal character_, in itself is a regex that matches itself. Thus, `a` matches third character in text `Charles`.  These literal characters are case sensitive.
Example-1
```{r}
ex_text <- "This is an example text"
# Match literal `x`
grepl(pattern = "x", ex_text)
# Match Upper case literal "X"
grepl("X", ex_text)
```

## Character sets
It is always not feasible to put every literal characters.  We may also match literal characters from a given set of options. To **match a group of characters** we have to put all these in square brackets.  So, `[abc]` matches either of `a`, `b`, or `c`.  
Example-
```{r}
ex_vec <- c("Apple", "Orange", "Myrrh")
# matches a vowel
grepl("[aeiou]", ex_vec)
```
To **match a range of characters/numbers** we can separate these by hyphen in square brackets. So, `[a-n]` will match a character from range `[abcdefghijklmn]`.
Example-
```{r}
ex_text <- "The quick brown fox jumps over the lazy dog"
grepl("[a-z]", ex_text)
grepl("[X-Z]", ex_text)
```
Example-2
```{r}
ex_colors <- c("grey", "black", "gray")
grepl("gr[ae]y", ex_colors)
```

We can also use __pre-built character classes__ listed below.

  + `[:punct:]` punctuation.
  + `[:alpha:]` letters.
  + `[:lower:]` lowercase letters.
  + `[:upper:]` uppercase letters.
  + `[:digit:]` digits.
  + `[:xdigit:]` hex digits.
  + `[:alnum:]` letters and numbers.
  + `[:cntrl:]` control characters.
  + `[:graph:]` letters, numbers, and punctuation.
  + `[:print:]` letters, numbers, punctuation, and white-space.
  + `[:space:]` space characters (basically equivalent to `\\s`).
  + `[:blank:]` space and tab.

Example-
```{r}
ex_vec2 <- c("One apple", "2 Oranges")
grepl("[:digit:]", ex_vec2)
```

## Non-printable characters/ Meta characters (short-hand character classes)

We can use special character sequences to put non-printable characters in our regular expression(s). E.g. `\t` matches a tab character.  __But since `\` is an escape character in R, we need to escape it too.__  So to match a tab character we have to put `\\t` in our regex sequence.  Regex for that matches new line (line feed) is `\\n`. `Regex` for other meta characters is listed below-

  + `\\s` matches a white-space character.  Moreover, its complement `\\S` matches any character except a white-space.
  + `\\w` matches any alphanumeric character. Similarly, its complement is `\\W` which matches any character except alphanumeric charachters.
  + `\\d` matches any digit.  Similarly, its complement is `\\D` which matches any character except digits.
  + `\\b` matches any word boundary.  Thus, `\\B` matches any character except a word boundary.
  + `.` matches any character. To match a literal dot `.` we have to escape that; and thus `\\.` matches a dot charcter.
  
See these examples-
```{r}
ex_vec3 <- c("One apple", "2 oranges & 3 bananas.")
# match word boundary
grepl("\\w", ex_vec3)
# match any character followed by a dot character
grepl(".\\.", ex_vec3)
```

## Quantifiers

What if we want to match more than one literal/character through `regex`? Let's say if we want to check whether the given string or string vector contain two consecutive vowels.  One method may be to use character classes two times i.e. using `[aeiou][aeiou]`. But this method is against the principles of __DRY__^[Dont repeat yourself] which is one of the common principle of programming.  To solve these issues, we have quantifiers.

- `+` __1 or more__ occurrences
- `*` __0 or more__
- `?` __0 or 1__
- `{}` specified numbers
  + `{n}` exactly n
  + `{n,}` n or more
  + `{n,m}` between n and m

Thus, we may match two consecutive vowels using `[aeiou]{2}`.  See this example

```{r}
ex_vec <- c("Apple", "Banana", "pineapple")
grepl("[aeiou]{2}", ex_vec)
```
## Alternation

  - `|` explain precedence

## Anchors

  - `^`
  - `$`

## Capture Groups

- `()`

## Lookaround

Lookahead and lookbehinds are zero-width assertions in regex. They are used to match a pattern only if it is followed or preceded by another pattern, respectively. The pattern in the lookahead or lookbehind is not included in the match.

### Lookahead

A lookahead is used to match a pattern only if it is followed by another pattern. *Positive Lookaheads* are written as `(?=...)`, where `...` is the pattern that must follow the match.

For example, the regex pattern `hello(?= world)` matches "hello" only if it is followed by " world". It matches "hello world" but not "hello there world" or "hello".

Example

```{r}
string <- c("hello world", "hello there world", "hello")
grep("hello(?= world)", string, value = TRUE, perl = TRUE)

```

### Lookbehind

A lookbehind is used to match a pattern only if it is preceded by another pattern. Lookbehinds are written as `(?<=...)`, where `...` is the pattern that must precede the match.

For example, the regex pattern `(?<=hello )world` matches "world" only if it is preceded by "hello ". It matches "hello world" but not "world hello" or "hello there world".

Example

```{r}
string <- c("hello world", "world hello", "hello there world")
grep("(?<=hello )world", string, value = TRUE, perl = TRUE)

```

### Negative Lookahead and Lookbehinds

Negative lookahead and negative lookbehinds are used to match a pattern only if it is not followed or preceded by another pattern, respectively. Negative lookahead and lookbehinds are written as `(?!...)` and `(?<!...)`, respectively.

For example, the regex pattern `hello(?! world)` matches "hello" only if it is not followed by " world". It matches "hello there" but not "hello world" or "hello world there".

Example-


```{r}
string <- c("hello there", "hello world", "hello world there")
grep("hello(?! world)", string, value = TRUE, perl = TRUE)

```

And the regex pattern `(?<!hello )world` matches "world" only if it is not preceded by "hello ". It matches "world hello" and "hello there world" but not "hello world".

```{r}
string <- c("hello world", "world hello", "hello there world")
grep("(?<!hello )world", string, value = TRUE, perl = TRUE)
```


## Comments

- `(?#..)`
- `regex(... , comments = TRUE)`


