--- 
title: "Data Analytics for audit using R"
author: "Anil Goyal"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This book is intended for auditors performing exploratory data analytics using R programming language, mainly baseR and Tidyverse."
---

```{r setup, include=FALSE}
options(
  htmltools.dir.version = FALSE, formatR.indent = 2, width = 55, digits = 4
)

# install the packages needed by this book; you fill out c(), e.g. c('ggplot2', 'dplyr')
lapply(c('xfun'), function(pkg) {
  if (system.file(package = pkg) == '') install.packages(pkg)
})


options(tinytex.verbose = TRUE)
```

# Preface {-}



## Why read this book {-}



## Structure of the book {-}




## Software information and conventions {-}

Packages **knitr**\index{knitr} [@xie2015] and the **bookdown**\index{bookdown} [@R-bookdown] have been used to compile this book. My R session information is shown below:

```{r}
xfun::session_info()
```

Package names are in bold text (e.g., **rmarkdown**), and inline code and filenames are formatted in a typewriter font (e.g., `knitr::knit('foo.Rmd')`). Function names are followed by parentheses (e.g., `bookdown::render_book()`).

## Acknowledgments {-}

A lot of people helped me when I was writing the book.

```{block2, type='flushright', html.tag='p'}
Anil Goyal
```


